---
title: "Chap 4. Fitting GARCH models"
EX: "Tương tự 4.16 ARMA (1,1) & GARCH(1,1)"
---
```{r}
library(xts)
library(rugarch)
library(qrmdata)
library(qrmtools)
doPDF <- require(crop)
options(digits = 10)
```

```{r}
## Dow Jones stock price data
data("DJ_const")
str(DJ_const)

## extract AAPL stock a time period and build log-returns
AAPLdata <- DJ_const['1990-01-01/1999-12-31',"AAPL"]
X <- -returns(AAPLdata)
```
### a) GARCH(1,1)
### Fit a GARCH(1,1) model with normal innovations 
```{r}
uspec.N <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                      mean.model = list(armaOrder = c(0,0 ), # AR(1) part
                                        include.mean = TRUE), # with mean
                      distribution.model = "norm") # standard normal innovations
(fit.N <- ugarchfit(spec = uspec.N, data = X)) # fit
```
###  Fit a GARCH(1,1) model with Student t innovations 
```{r}
uspec.t <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                      mean.model = list(armaOrder = c(0,0), include.mean = TRUE),
                      distribution.model = "std") # standardized Student t innovations
(fit.t <- ugarchfit(spec = uspec.t, data = X)) # fit
```
###  Fit a GARCH(1,1) model with Skewed Student t innovations (Skewed Student t distribution - sstd)
```{r}
uspec.st <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                      mean.model = list(armaOrder = c(0,0), include.mean = TRUE),
                      distribution.model = "sstd")
(fit.st <- ugarchfit(spec = uspec.st, data = X)) # fit
```
### b) ARMA (1,1) + GARCH(1,1)
```{r}
uspec.N1 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                      mean.model = list(armaOrder = c(1,1), # AR(1) part
                                        include.mean = TRUE), # with mean
                      distribution.model = "norm") # standard normal innovations
(fit.N1 <- ugarchfit(spec = uspec.N1, data = X)) # fit
```
### Fit a ARMA (1,1) + GARCH(1,1) model with Student t innovations (Student t distribution - std)
```{r}
uspec.t1 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                      mean.model = list(armaOrder = c(1,1), include.mean = TRUE),
                      distribution.model = "std") # standardized Student t innovations
(fit.t1 <- ugarchfit(spec = uspec.t1, data = X)) # fit
```
### Fit a ARMA (1,1) + GARCH(1,1) model with with Skewed Student t innovations (Skewed Student t distribution - sstd)
```{r}
uspec.st1 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                      mean.model = list(armaOrder = c(1,1), include.mean = TRUE),
                      distribution.model = "sstd")
(fit.st1 <- ugarchfit(spec = uspec.st1, data = X)) # fit

```

### Comparing 2 models  
###  Likelihood-ratio test for comparing the two models ##
```{r}
## The likelihood ratio statistic is twice the difference between the log-likelihoods

LL.N <- fit.N@fit$LLH # log-likelihood of the model based on normal innovations
LL.N1 <- fit.N1@fit$LLH # log-likelihood of the model based on normal innovations

LL.t <- fit.t@fit$LLH # log-likelihood of the model based on t innovations
LL.t1 <- fit.t1@fit$LLH # log-likelihood of the model based on t innovations

LL.st <- fit.st@fit$LLH # log-likelihood of the model based on t innovations
LL.st1 <- fit.st1@fit$LLH # log-likelihood of the model based on t innovations
```
###  Using Akaike's information criterion to compare the two models
```{r}
(AIC.N <- 2*length(fit.N@fit$coef) - 2*LL.N)
(AIC.N1 <- 2*length(fit.N1@fit$coef) - 2*LL.N1)

```

```{r}
(AIC.t <- 2*length(fit.t@fit$coef) - 2*LL.t)
(AIC.t1 <- 2*length(fit.t1@fit$coef) - 2*LL.t1)

```

```{r}
(AIC.st <- 2*length(fit.st@fit$coef) - 2*LL.st)
(AIC.st1 <- 2*length(fit.st1@fit$coef) - 2*LL.st1)

```

### c) VaR
## VaR chung (kco ngày cụ thể)
```{r}
VaR.99 <- quantile(fit.t, probs = 0.99) # estimated VaR at level 99%
```

```{r}
tail(VaR.99)
```

```{r}
VaR.99['1999-12-31']
```

### d) GARCH volatility estimate
## EWMA estimate of volatility
```{r}
uspec.EWMA <- ugarchspec(variance.model = list(model = "iGARCH", garchOrder = c(1, 1)),
                         mean.model = list(armaOrder = c(0, 0), include.mean = TRUE))

fit <- ugarchfit(spec = uspec.EWMA, data = X)  # fit
vola.EWMA <- as.numeric(sigma(fit)) # extract vola; in-sample (sigma_t)

```

```{r}
vola = sigma(fit)
```
## Comparison with the GARCH volatility estimate
```{r}
time <- index(vola)
plot(time, as.numeric(vola), type = "l",
     xlab = "Time t", ylab = expression("Estimated volatility"~(sigma[t])))
lines(time, vola.EWMA, col = "royalblue3")
legend("topleft", bty = "n", lty = rep(1, 2), col = c("black", "royalblue3"),
       legend = c("GARCH(1,1) volatility estimate", "EWMA estimate of volatility"))

## Relative difference
plot(time, abs(vola - vola.EWMA)/abs(vola), type = "l", xlab = "Time t",
     ylab = "Relative difference between the two volatility estimates")

```
## Summary statistics of the relative differencea => The relative difference between the two estimated volatilities
```{r}
summary(as.numeric(abs(vola - vola.EWMA)/abs(vola)))
```

