---
title: "09_Standard_methods_for_market_risk & 9.18 Standard methods for market risk"
---
```{r}
library(nvmix) # for rNorm(), fitStudent(), rStudent()
library(xts) # for na.fill()
library(qrmdata) # for the data
library(qrmtools) # for the data analysis
```

## 09_Standard_methods_for_market_risk
### 1 Working with the (BMW, Siemens) data
```{r}
data(EURSTX_const)
db <- EURSTX_const
```

```{r}
## Pick out the sub-database of stock prices and fill gaps
time <- c("2000-01-03", "2009-12-31")
S <- db[paste0(time, collapse = "/"), c("BMW.DE", "SIE.DE")] # pick out data (data.frame)
str(S)
S <- na.omit(S) # fill NAs 
```

## Compute the risk-factor changes
```{r}
X <- returns(S) # compute log-returns
X <- as.matrix(X)
```

### 2 Implement (and document) some auxiliary functions
```{r}
loss_operator <- function(x, weights)
    -rowSums(expm1(x) * matrix(weights, nrow = nrow(x), ncol = length(weights), byrow = TRUE))

risk_measure <- function(S, lambda, alpha,
                         method = c("Var.Cov", "hist.sim"),
                         ...)
{
    ## Input checks and conversions
    if(!is.matrix(S)) S <- rbind(S, deparse.level = 0L) 
    stopifnot(0 < alpha, alpha < 1, 
              
              length(lambda) == ncol(S), lambda > 0) 
    
    method <- match.arg(method) 
    

    ## Ingredients required for *all* methods
    X <- returns(S) # compute risk-factor changes
    if(!length(X)) stop("'S' should have more than just one line")
    S. <- as.numeric(tail(S, n = 1)) 
    w. <- lambda * S. 

    ## Method switch 
    switch(method,
           
# variance-covariance method
           "Var.Cov" = { 
               mu.hat <- colMeans(X) 
               Sigma.hat <- var(X) 
               L.delta.mean <- -sum(w. * mu.hat) 
               L.delta.sd <- sqrt(t(w.) %*% Sigma.hat %*% w.) 
## Compute VaR and ES and return
               qa <- qnorm(alpha)
               list(VaR = L.delta.mean + L.delta.sd * qa,
                    ES  = L.delta.mean + L.delta.sd * dnorm(qa) / (1-alpha),
                    mu    = mu.hat, 
                    Sigma = Sigma.hat) 
           },

# historical simulation method
           "hist.sim" = { 
               L <- loss_operator(X, weights = w.) 
               list(VaR = VaR_np(L, alpha),
                    ES  =  ES_np(L, alpha))
           },
         stop("Wrong 'method'")
    )
}
```

### 3 Compute VaR and ES for the standard methods
### thầy cho lamda và alpha, thay số phần này
```{r}
lambda <- c(1, 10) # thầy sẽ cho nhớ thay
alpha <- 0.99
```

## Estimate VaR and ES with the various methods
```{r}
var.cov  <- risk_measure(S, lambda = lambda, alpha = alpha, method = "Var.Cov")
hist.sim <- risk_measure(S, lambda = lambda, alpha = alpha, method = "hist.sim")
```

## Pick out VaR and ES for all methods
```{r}
(rm <- rbind("Var.-cov."      = unlist(var.cov[c("VaR", "ES")]),
             "Hist. sim."     = unlist(hist.sim[c("VaR", "ES")])))
# [Giải thích cho output]
# Var.-cov
# VaR ≈ 33.06: on about 99% of days, the 1-day loss is expected to be no more than ~33.06; about 1% of days the loss can be larger.
# ES ≈ 37.87: conditional on being in the worst 1% of days (i.e., beyond VaR), the average 1-day loss is about 37.87.

# Hist. sim
# VaR ≈ 38.35: based on the historical loss distribution, about 1% of days can have losses exceeding ~38.35.
# ES ≈ 56.27: within the worst 1% of days, the average loss rises to about 56.27.
```
### 4 Graphical analysis
```{r}
## Compute historical losses 
S. <- as.numeric(tail(S, n = 1)) # pick out last available stock prices ("today")
w. <- lambda * S. # weights tilde{w}
L <- loss_operator(X, weights = w.) # historical losses
summary(L) # get important statistics about the losses
```

#___________
### 9.18 Standard methods for market risk
```{r}
install.packages("sfsmisc")   # cách chuẩn
library(RColorBrewer)
library(rugarch)
library(sfsmisc) # for eaxis()
library(nvmix) # for rNorm(), fitStudent() and rStudent()
library(zoo) # for na.fill()
library(xts) # for xts()
library(ADGofTest) # for ad.test()
library(qrmdata)
library(qrmtools)
doPDF <- require(crop)
options(digits = 10)
```

## a), b) Compute the risk-factor changes
## Data preparation
```{r}
data(EURSTX_const)
time <- c("2000-01-03", "2009-12-31")
S <- EURSTX_const[paste0(time, collapse = "/"), c("BMW.DE", "SIE.DE")] 
colnames(S) <- c("BMW", "SIE")
S <- na.omit(S)
```

## Compute the risk-factor changes
```{r}
X <- returns(S)
X <- as.matrix(X)
```

## c) Ingredients required for *all* methods
```{r}
loss_operator <- function(x, weights)
    -rowSums(expm1(x) * matrix(weights, nrow = nrow(x), ncol = length(weights), byrow = TRUE))

risk_measure <- function(S, lambda, alpha,
                         method = c("Var.Cov", "hist.sim", "MC.N", "MC.t", "POT"),
                         ...)
{
    ## Input checks and conversions
    if(!is.matrix(S)) S <- rbind(S, deparse.level = 0L) 
    stopifnot(0 < alpha, alpha < 1,
              length(lambda) == ncol(S), lambda > 0) 
    method <- match.arg(method) 

    ## Ingredients required for *all* methods
    X <- returns(S) 
    if(!length(X)) stop("'S' should have more than just one line") 
    S. <- as.numeric(tail(S, n = 1)) 
    w. <- lambda * S. 

    ## Method switch 
    switch(method,
           
# variance-covariance method           
           "Var.Cov" = { 
               mu <- colMeans(X) 
               Sigma <- var(X) 
               L.Delta.mean <- -sum(w. * mu)
               L.Delta.sd <- sqrt(t(w.) %*% Sigma %*% w.) 

## Compute VaR and ES and return
               qa <- qnorm(alpha)
               list(VaR = L.Delta.mean + L.Delta.sd * qa,
                    ES  = L.Delta.mean + L.Delta.sd * dnorm(qa) / (1-alpha),
                    mu    = mu, 
                    Sigma = Sigma) 
           },

# historical simulation method
           "hist.sim" = { 
               L <- loss_operator(X, weights = w.) 
               list(VaR = VaR_np(L, alpha),
                    ES  =  ES_np(L, alpha))
           },
# Monte Carlo based on a fitted multivariate normal
           "MC.N" = { 
               stopifnot(hasArg(N)) 
               N <- list(...)$N 
               mu <- colMeans(X) 
               Sigma  <- var(X) 
               X. <- rNorm(N, loc = mu, scale = Sigma) 
               L <- loss_operator(X., weights = w.) 

## Compute VaR and ES and return
               list(VaR = VaR_np(L, alpha), 
                    ES  =  ES_np(L, alpha), 
                    mu    = mu,
                    Sigma = Sigma) 
           },

# Monte Carlo based on a fitted multivariate t
           "MC.t" = { 
               stopifnot(hasArg(N)) 
               N <- list(...)$N 
               fit <- fitStudent(X) 
               X. <- rStudent(N, df = fit$df, loc = fit$loc, scale = fit$scale) 
               L <- loss_operator(X., weights = w.) 

## Compute VaR and ES and return
               list(VaR = VaR_np(L, alpha), 
                    ES  =  ES_np(L, alpha), 
                    mu    = fit$loc, 
                    sigma = fit$scale, 
                    df    = fit$df) 
           },

 # Generalized Pareto distribution (GPD)
           "POT" = { 
               stopifnot(hasArg(q)) 
               L. <- loss_operator(X, weights = w.)
               u <- quantile(L., probs = list(...)$q, names = FALSE) 
               excess <- L.[L. > u] - u
               fit <- fit_GPD_MLE(excess) 
               xi <- fit$par[["shape"]] 
               beta <- fit$par[["scale"]]
               if(xi <= 0) stop("Risk measures only implemented for xi > 0.")
               
## Compute semi-parametric VaR and ES estimates
               Fbu <- length(excess) / length(L.) 
               VaR <- u + (beta/xi)*(((1-alpha)/Fbu)^(-xi)-1) 
               ES <- (VaR + beta-xi*u) / (1-xi) 
               if(xi >= 1) ES <- Inf 
               
## Return
               list(VaR = VaR, 
                    ES  = ES, 
                    xi  = xi, 
                    beta = beta, 
                    converged = fit$converged, 
                    u = u, 
                    excess = excess) 
           },
           stop("Wrong 'method'"))
}
```

## d) VaR and ES for all methods
```{r}
lambda <- c(1, 10)
alpha <- 0.99 
N <- 1e4 # Monte Carlo sample size

## Estimate VaR and ES with the various methods
set.seed(271) 

var.cov  <- risk_measure(S, lambda = lambda, alpha = alpha, method = "Var.Cov")
hist.sim <- risk_measure(S, lambda = lambda, alpha = alpha, method = "hist.sim")
MC.N     <- risk_measure(S, lambda = lambda, alpha = alpha, method = "MC.N", N = N)
POT      <- risk_measure(S, lambda = lambda, alpha = alpha, method = "POT",  N = N, q = 0.9)
MC.t     <- risk_measure(S, lambda = lambda, alpha = alpha, method = "MC.t", N = N)

## Pick out VaR and ES for all methods
rm <- rbind("MC (normal)"    = unlist(MC.N[c("VaR", "ES")]),
            "Var.-cov."      = unlist(var.cov[c("VaR", "ES")]),
            "Hist. sim."     = unlist(hist.sim[c("VaR", "ES")]),
            "POT"            = unlist(POT [c("VaR", "ES")]),
            "MC (Student t)" = unlist(MC.t[c("VaR", "ES")]))
round(rm, digits = 4)
```

## e) Compute historical losses 
```{r}
S. <- as.numeric(tail(S, n = 1)) 
w. <- lambda * S. 
L <- loss_operator(X, weights = w.) 
summary(L) # get important statistics about the losses
```

#___________
### Bài thầy chữa
```{r}
# ĐỀ BÀI
# DJ [2006-01-03/2016-31-12]
# AAPL, IBM, INTC
# lambda 2 1 2
```

## a), b) Compute the risk-factor changes
## Data preparation
```{r}
data(DJ_const)
time <- c("2006-01-03", "2016-12-31")
S <- DJ_const[paste0(time, collapse = "/"), c("AAPL", "IBM", "INTC")] # pick out data (data.frame)
colnames(S) <- c("AAPL", "IBM", "INTC")
S <- na.omit(S)  
```

## c) Ingredients required for *"Var.Cov"& "hist.sim"* methods
```{r}
loss_operator <- function(x, weights)
    -rowSums(expm1(x) * matrix(weights, nrow = nrow(x), ncol = length(weights), byrow = TRUE))

risk_measure <- function(S, lambda, alpha,
                         method = c("Var.Cov", "hist.sim"),
                         ...)
{
    ## Input checks and conversions
    if(!is.matrix(S)) S <- rbind(S, deparse.level = 0L) 
    stopifnot(0 < alpha, alpha < 1,
              length(lambda) == ncol(S), lambda > 0) 
    method <- match.arg(method) 

    ## Ingredients required for *all* methods
    X <- returns(S) 
    if(!length(X)) stop("'S' should have more than just one line") 
    S. <- as.numeric(tail(S, n = 1)) 
    w. <- lambda * S. 

    ## Method switch 
    switch(method,
           
# variance-covariance method           
           "Var.Cov" = { 
               mu <- colMeans(X) 
               Sigma <- var(X) 
               L.Delta.mean <- -sum(w. * mu)
               L.Delta.sd <- sqrt(t(w.) %*% Sigma %*% w.) 

## Compute VaR and ES and return
               qa <- qnorm(alpha)
               list(VaR = L.Delta.mean + L.Delta.sd * qa,
                    ES  = L.Delta.mean + L.Delta.sd * dnorm(qa) / (1-alpha),
                    mu    = mu, 
                    Sigma = Sigma) 
           },

# historical simulation method
           "hist.sim" = { 
               L <- loss_operator(X, weights = w.) 
               list(VaR = VaR_np(L, alpha),
                    ES  =  ES_np(L, alpha))
           },
           stop("Wrong 'method'"))
}
```

## d) VaR and ES *"Var.Cov"& "hist.sim"* methods
```{r}
lambda <- c(2,1,2) # thầy cho
alpha <- 0.99 # confidence levels for computing the risk measures

## Estimate VaR and ES with the various methods
var.cov  <- risk_measure(S, lambda = lambda, alpha = alpha, method = "Var.Cov")

hist.sim <- risk_measure(S, lambda = lambda, alpha = alpha, method = "hist.sim")

## Pick out VaR and ES for all methods
rm <- rbind("Var.-cov."      = unlist(var.cov[c("VaR", "ES")]),
            
            "Hist. sim."     = unlist(hist.sim[c("VaR", "ES")]))
round(rm, digits = 4)
```
`

