---
title: "E5. Practice - Black Maxima"
---
```{r}
library(xts)
library(qrmdata)
library(qrmtools)
doPDF <- require(crop)
options(digits = 10)
```

```{r}
profLogLik <- function(r, k, x, control = list(reltol = 0), ...)
{
    ## log-likelihood as a function of the nuisance parameters xi, sigma
    ## (mu is implied from the given r, k)
    ## Note: Could very well be -Inf due to choice of r or k and thus implied mu
    logLik.xi.sig <- function(theta) { # note: r, k and x are locally 'seen' here
        xi <- theta[1]
        sig <- theta[2]
        imu <- if(xi == 0) { # mu implied by xi, sigma, r, k
                   r + sig * log(-log1p(-1/k))
               } else {
                   r - (sig/xi) * ((-log1p(-1/k))^(-xi) - 1)
               }
        stopifnot(is.finite(imu)) # sanity check
        sum(dGEV(x, shape = xi, loc = imu, scale = sig, log = TRUE)) # log-likelihood (correctly deals with sigma <= 0)
    }
    

    init <- c(0, sqrt(6 * var(x)) / pi) # case 'xi = 0' (xi, sigma)
    while(!is.finite(logLik.xi.sig(init))) init[2] <- init[2] * 2
    ## Maximize the log-likelihood in the nuisance parameters theta = (xi, sigma)
    ## for our given parameters of interest r and k (and the data x)
    control <- c(as.list(control), fnscale = -1) # maximization (overwrites possible additionally passed 'fnscale')
    optim(init, fn = logLik.xi.sig, control = control, ...) # maximize logLik.xi.sig
}


obj <- function(r, k, x, maxLogLik, alpha = 0.05, control = list(reltol = 0), ...)
    (profLogLik(r = r, k = k, x = x, control = control, ...)$value) - # profile log-likelihood
        (maxLogLik - qchisq(1-alpha, df = 1) / 2)

plot_obj <- function(r, k, x, maxLogLik, alpha = 0.05, control = list(reltol = 0), ...)
{
    lr <- length(r)
    lk <- length(k)
    stopifnot(xor(lr == 1, lk == 1)) # one of them has to have length > 1 (otherwise no plot)
    if(lr > 1) {
        x. <- r
        y. <- sapply(r, function(r.) obj(r = r., k = k, x = x,
                                         maxLogLik = maxLogLik, alpha = alpha,
                                         control = control, ...))
        xlab <- "Return level r"

    } else { # length(k) > 1
        x. <- k
        y. <- sapply(k, function(k.) obj(r = r, k = k., x = x,
                                         maxLogLik = maxLogLik, alpha = alpha,
                                         control = control, ...))
        xlab <- "Return period k"
    }
    plot(x., y., type = "l", xlab = xlab,
         ylab = "Objective function for finding root")
    abline(h = 0, lty = 2)
}
```

```{r}
data("DJ_const")
str(DJ_const)

DJdata <- DJ_const['1986-01-01/2006-12-31', 'IBM']
#IBM <- na.fill(DJdata, fill = "extend")
IBMreturn <- -returns(DJdata)
plot.zoo(IBMreturn, xlab = "Time")
```

```{r}
## Extract (half-)yearly maxima
M.y <- period.apply(IBMreturn, INDEX = endpoints(IBMreturn, "years"), FUN = max) # yearly maxima
endpts <- endpoints(IBMreturn, "quarters") # end indices for quarters
endpts <- endpts[seq(1, length(endpts), by = 2)] # end indices for half-years
M.hy <- period.apply(IBMreturn, INDEX = endpts, FUN = max) # half-yearly maxima
dim(M.y)
```

```{r}
head(M.y)
tail(M.y,21)
```

```{r}
#để xóa dữ liệu NaN
M.y <- na.omit(M.y)
dim(M.y)
```

```{r}
length(M.y)
anyNA(M.y)
any(!is.finite(coredata(M.y)))
coredata(M.y)[!is.finite(coredata(M.y))]
```
## Fit GEV to yearly maxima
```{r}
fit.y <- fit_GEV_MLE(M.y) # maximum likelihood estimator
(xi.y <- fit.y$par[["shape"]]) # ~= 0.2971 => Frechet domain with infinite ceiling(1/xi.y) = 4th moment
(mu.y  <- fit.y$par[["loc"]])
(sig.y <- fit.y$par[["scale"]])
fit.y$SE # standard errors
mLL <- fit.y$value # maximum log-likelihood (at MLE)
```
## Fix the return period k = 20 (in years) and estimate the corresponding
## return level r
```{r}
k <- 20 # fix return period
r <- qGEV(1-1/k, shape = xi.y, loc = mu.y, scale = sig.y) # corresponding estimated return level r
r
```
## Find a suitable initial interval for computing the CI
```{r}
I <- c(0.1, 0.5) # found by experimenting with the following plot (Điều chỉnh các giá trị để đường =0 cắt plot tại 2 điểm)
plot_obj(r = seq(I[1], I[2], length.out = 128), k = k, x = M.y, maxLogLik = mLL)
abline(v = r) # estimated return level
```
## Compute 95%-CI for the 10-year return level r
```{r}
CI.low <- uniroot(obj, lower = I[1], upper = r,    k = k, x = M.y, maxLogLik = mLL)
CI.low
```
```{r}
CI.up  <- uniroot(obj, lower = r,    upper = I[2], k = k, x = M.y, maxLogLik = mLL)
CI.up
```

```{r}
(CI.r10 <- c(CI.low$root, CI.up$root)) # 95%-CI for 20-year return level r; [0.137, 0.412]
```
### 2.1.2 CI for the return period of a loss as on Black Monday
## Fix the return level r (as on Black Monday) and estimate the corresponding
## return period k
```{r}
r <- 0.65 # fix return level
k <- 1/(1-pGEV(r, shape = xi.y, loc = mu.y, scale = sig.y)) # corresponding estimated return period k
k
```
## Find a suitable initial interval for computing the CI
```{r}
I <- c(30,100) # found by experimenting with the following plot
plot_obj(r = r, k = seq(I[1], I[2], length.out = 128), x = M.y, maxLogLik = mLL)
abline(v = k) # estimated return level
```

```{r}
CI.low <- uniroot(obj, lower = I[1], upper = k, r = r, x = M.y, maxLogLik = mLL)
CI.low$root
```

